High-level cognition is supported by information-rich but compressible brain activity patterns

### 2023-12-20 paper/main.tex
working on introduction text related to fig 1.

To the extent that brain activity patterns contain rich task-relevant
information, we should be able to use the activity patterns to accurately
differentiate between different aspects of a task~\citep[e.g., using pattern
classifiers;][]{NormEtal06b}. For example, prior work has shown a direct
correspondence between classification accuracy and the information content of a
signal~\citep{Alva02}. To the extent that brain activity patterns are
compressible, we should be able to generate simplified (e.g., lower
dimensional) representations of the data while still preserving the relevant or
important aspects of the original signal. In general, information content and
compressibility are related but are partially dissociable
(Fig.~\ref{fig:information-compression}). If a given signal (e.g., a
representation of brain activity patterns) contains more information about
ongoing cognitive processes, then the peak decoding accuracy should be high.
And if the signal is compressible, a low-dimensional embedding of the signal
will be similarly informative as the original signal
(Fig.~\ref{fig:information-compression}D).

@@

>>>>

To the extent that brain activity patterns contain rich task-relevant
information, we should be able to use the activity patterns to accurately
differentiate between different aspects of a task~\citep[e.g., using pattern
classifiers;][]{NormEtal06b}. For example, prior work has shown a direct
correspondence between classification accuracy and the information content of a
signal~\citep{Alva02}. To the extent that brain activity patterns are
compressible, we should be able to generate simplified (e.g., lower
dimensional) representations of the data while still preserving the relevant or
important aspects of the original signal. In general, information content and
compressibility are often related but are also dissociable
(Fig.~\ref{fig:information-compression}). If a given signal (e.g., a
representation of brain activity patterns) contains more information about
ongoing cognitive processes, then the peak decoding accuracy should be high. In
the simulations shown in Figure~\ref{fig:information-compression}C we construct
synthetic datasets that have high or low levels of informativeness by varying
temporal autocorrelations in the data (see \textit{Synthetic data}).
If a signal is compressible, then a low-dimensional embedding of the signal
will be similarly informative as the original signal.  In the simulations
shown in Figure~\ref{fig:information-compression}C we construct synthetic
datasets that have high or low levels of compressibility by varying
the covariance structure across features (see \textit{Synthetic data}).  As shown
in Figure~\ref{fig:information-compression}D, highly informative datasets
yield higher decoding accuracies than less informative datasets (i.e., the peaks of the curves in
the top panels of Fig.~\ref{fig:information-compression}D are higher than the
peaks of the curves in the bottom panels). Highly compressible datasets show
steeper slopes when we plot decoding accuracy as a function of the number of
components used to represent the data (i.e., the slopes of the curves in the
left panels of Fig.~\ref{fig:information-compression}D are steeper than the
slopes of the curves in the right panels).  Whereas characterizing
the informativeness and compressibility of synthetic data can be instructive,
we are ultimately interested in understanding how these properties relate to
brain activity patterns recorded under different cognitive circumstances.

@@

Unless otherwise noted, all statistical tests are two-sided, all error bars and
error ribbons denote bootstrap-estimated 95\% confidence intervals across
participants, and all reported $p$-values are corrected for multiple
comparisons using the Benjamini-Hochberg procedure~\citep{BenjHoch95}.
